---
title: "Land owned by faith-based institutions"
author: ""
output:
  html_document:
    number_sections: FALSE
    self_contained: TRUE
    code_folding: show
    toc: TRUE
    toc_float: TRUE
    editor_options:
      chunk_output_type: console
---

<link rel="stylesheet" href="//fonts.googleapis.com/css?family=Lato" />

Library: Requests

Author: Sarah Strochak

Version: R 3.6.1

Last updated `r format(Sys.time(), '%B %d, %Y')`

Environment: Local Windows session (desktop)

```{r rmarkdown-setup, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r setup}
options(scipen = 999)

library(tidyverse)
```

## Read in church-owned parcels

```{r read-churches}

churches <- read_csv("L:/Libraries/Requests/Raw/2019/dmv_who_owns_church.csv") %>% 
  mutate(fipscodestatecounty = as.character(fipscodestatecounty))

# Merge on county names
churches <- urbnmapr::get_urbn_map(map = "counties", sf = TRUE) %>% 
  sf::st_drop_geometry() %>% 
  select(county_name, county_fips) %>% 
  right_join(churches, by = c("county_fips" = "fipscodestatecounty"))

```

Subset Arlington.

```{r subset}

arl <- churches %>% 
  filter(county_name == "Arlington County")
  

```

## Read in zoning codes

For each county, choose only one record per zoning code. It only matters if residential is allowed, not the specific speculations of the zoning code.

```{r read-zoning}

arl_zoning <- read_csv("L:/Libraries/RegHsg/Doc/zoning-codes/arlington-zoning-code.csv") %>% 
  mutate(county_name = "Arlington County",
         residential = 1) %>% 
  select(zoning_code = Code, county_name, residential) %>% 
  distinct()

ff_zoning <- read_csv("L:/Libraries/RegHsg/Doc/zoning-codes/fairfax-zoning-code.csv") %>% 
  mutate(county_name = "Fairfax County",
         residential = 1) %>% 
  rename(zoning_code = code) %>% 
  select(zoning_code, county_name, residential) %>% 
  distinct()

dc_zoning <- read_csv("L:/Libraries/RegHsg/Doc/zoning-codes/DC Zoning code.csv") %>% 
  mutate(county_name = "District of Columbia",
         residential = 1) %>% 
  rename(zoning_code = Code) %>% 
  select(zoning_code, county_name, residential) %>% 
  distinct()

```


## Arlington County

### Merge on zoning codes


```{r merge-zoning}

arl <- arl %>% 
  mutate(zoning_code = case_when(
    zoning == "RC" ~ "R-C",
    zoning == "C-0 Crystal City" ~ "C-O Crystal City",
    zoning %in% c("C-O-ROSSLYN", "C-O-ROSS") ~ "C-O Rosslyn",
    zoning == "RA-H-3.2" ~ "RA-H 3.2",
    zoning == "C-1-O" ~ "C-1-0",
    TRUE ~ zoning
  ))

arl1 <- arl %>% 
  left_join(arl_zoning, by = c("zoning_code", "county_name")) %>% 
  mutate(residential = ifelse(is.na(residential),
                              0,
                              residential))
```

Count which zoning codes are non-residential.

```{r}

arl1 %>% 
  filter(residential == 0) %>% 
  count(zoning_code)
  
```

Since all of these are dual zoning codes, reclassify to residential.

```{r}
arl1 <- arl1 %>% 
  mutate(residential = ifelse(residential == 0,
                              1,
                              residential))
```

### Classify vacant vs. non-vacant


```{r}

arl1 %>% 
  group_by(countylandusedescription) %>% 
  summarize(parcels = n(),
            lotsize_missing = sum(is.na(lotsizeorarea)),
            lotsize_0 = sum(lotsizeorarea == 0),
            building_missing = sum(is.na(buildingarea)),
            building_0 = sum(buildingarea == 0),
            share_buildingarea_0 = building_0 / parcels) %>% 
  arrange(desc(share_buildingarea_0))

vac_codes <- c("VACANT COMMERCIAL",
               "VACANT RESIDENTIAL",
               "RESIDENTIAL COST-VAL - VAC/SF & TW",
               "GEN COMM VAC-LAND/NO SITE PLAN")

arl1 <- arl1 %>% 
  mutate(vacant = ifelse(countylandusedescription %in% vac_codes,
                         "Vacant",
                         "Non-vacant"))

arl1 %>% 
  group_by(vacant) %>% 
  summarize(parcels = n(),
            lotsize_missing = sum(is.na(lotsizeorarea)),
            lotsize_0 = sum(lotsizeorarea == 0),
            building_missing = sum(is.na(buildingarea)),
            building_0 = sum(buildingarea == 0))
```

### Calculation

Calculate land owned by faith-based institutions in Arlington.

```{r}

a1 <- arl1 %>% 
  mutate(lotsize_sqfeet = ifelse(lotsizeareaunit == "SF",
                               lotsizeorarea,
                               lotsizeorarea * 43560),
         free_site = lotsize_sqfeet - buildingarea,
         units = floor(((free_site * .75) / 950)) * 2) %>% 
  group_by(county_name, residential) %>% 
  summarize(parcels = n(),
            average_lot = mean(lotsize_sqfeet, na.rm = TRUE),
            average_empty = mean(lotsize_sqfeet[buildingarea == 0]),
            average_unused = mean(free_site, na.rm = TRUE),
            total_lot = sum(lotsize_sqfeet, na.rm = TRUE),
            empty_lot = sum(lotsize_sqfeet[buildingarea == 0]),
            unused_lot = sum(free_site, na.rm = TRUE),
            unused_over_1000 = sum(free_site > 1000, na.rm = TRUE),
            amount_over_1000 = sum(free_site[free_site > 1000], na.rm = TRUE),
            avg_units = mean(units, na.rm = TRUE),
            units = sum(units, na.rm = TRUE)) %>% 
  mutate(residential = "Total (all parcels allow residential") %>% 
  select(county_name, residential, everything())


```


## All other jurisdictions

Test variables 
```{r}

churches %>% 
  filter(county_name != "Arlington County") %>% 
  mutate(residential = ifelse(str_sub(zoning, 1, 1) == "R",
                              1,
                              0)) %>% 
  group_by(county_name, residential) %>% 
  summarize(parcels = n(),
            lotsize_missing = sum(is.na(lotsizeorarea)),
            lotsize_0 = sum(lotsizeorarea == 0),
            building_missing = sum(is.na(buildingarea)),
            building_0 = sum(buildingarea == 0))


```

```{r}


churches1 <- churches %>% 
  mutate(buildingarea = ifelse(is.na(buildingarea),
                               0,
                               buildingarea),
         vacant = ifelse(str_detect(countylandusedescription, "VACANT") == TRUE,
                         1,
                         0),
         vacant = ifelse(str_detect(countylandusedescription, "PARKING") == TRUE,
                         1,
                         vacant)) %>% 
  mutate(lotsize_sqfeet = ifelse(lotsizeareaunit == "SF",
                                 lotsizeorarea,
                                 lotsizeorarea * 43560),
         free_site = lotsize_sqfeet - buildingarea,
         free_site = ifelse(free_site < 0,
                            0,
                            free_site),
         residential = ifelse(str_sub(zoning, 1, 1) == "R",
                              "Residential zone",
                              "Non-residential zone"),
         residential = ifelse(is.na(zoning),
                              "Missing zoning information",
                              residential),
         units_base = floor((free_site * .75) / 950))


calc_units <- function(floors = 1,
                       sq_feet = 950) {
  
  totals <- churches1 %>% 
  mutate(units = units_base * floors) %>% 
  group_by(county_name) %>% 
  summarize(parcels = n(),
            empty_parcels = sum(buildingarea == 0, na.rm = TRUE),
            average_empty = mean(lotsize_sqfeet[buildingarea == 0], na.rm = TRUE),
            total_empty_lot = sum(lotsize_sqfeet[buildingarea == 0], na.rm = TRUE),
            avg_units = mean(units[buildingarea == 0], na.rm = TRUE),
            units = sum(units[buildingarea == 0], na.rm = TRUE)) %>% 
  mutate(residential = "Total")


all <- churches1 %>% 
  mutate(units = units_base * floors) %>% 
  group_by(county_name, residential) %>% 
  summarize(parcels = n(),
            empty_parcels = sum(buildingarea == 0, na.rm = TRUE),
            average_empty = mean(lotsize_sqfeet[buildingarea == 0], na.rm = TRUE),
            total_empty_lot = sum(lotsize_sqfeet[buildingarea == 0], na.rm = TRUE),
            avg_units = mean(units[buildingarea == 0], na.rm = TRUE),
            units = sum(units[buildingarea == 0], na.rm = TRUE)) %>% 
  bind_rows(totals) %>% 
  arrange(county_name, residential)


write_csv(all, here::here("Data", str_glue("church-lots_{floors}floors_empty.csv")))
  

}

walk(c(1, 2, 4, 5), ~ calc_units(floors = .))



calc_units_vacant <- function(floors = 1,
                       sq_feet = 950) {
  
  
  totals <- churches1 %>% 
    filter(vacant == 1) %>% 
  mutate(units = units_base * floors) %>% 
  group_by(county_name) %>% 
  summarize(parcels = n(),
            empty_parcels = sum(buildingarea == 0, na.rm = TRUE),
            average_empty = mean(lotsize_sqfeet[buildingarea == 0], na.rm = TRUE),
            total_empty_lot = sum(lotsize_sqfeet[buildingarea == 0], na.rm = TRUE),
            avg_units = mean(units[buildingarea == 0], na.rm = TRUE),
            units = sum(units[buildingarea == 0], na.rm = TRUE)) %>% 
  mutate(residential = "Total")


all <- churches1 %>% 
      filter(vacant == 1) %>% 
  mutate(units = units_base * floors) %>% 
  group_by(county_name, residential) %>% 
  summarize(parcels = n(),
            empty_parcels = sum(buildingarea == 0, na.rm = TRUE),
            average_empty = mean(lotsize_sqfeet[buildingarea == 0], na.rm = TRUE),
            total_empty_lot = sum(lotsize_sqfeet[buildingarea == 0], na.rm = TRUE),
            avg_units = mean(units[buildingarea == 0], na.rm = TRUE),
            units = sum(units[buildingarea == 0], na.rm = TRUE)) %>% 
  bind_rows(totals) %>% 
  arrange(county_name, residential)


write_csv(all, here::here("Data", 
                          str_glue("church-lots_{floors}floors_empty_vacant.csv")))
  

}

walk(c(1, 2, 4, 5), ~ calc_units(floors = .))




```




