---
title: "Land owned by faith-based institutions"
author: ""
output:
  html_document:
    number_sections: FALSE
    self_contained: TRUE
    code_folding: show
    toc: TRUE
    toc_float: TRUE
    editor_options:
      chunk_output_type: console
---

<link rel="stylesheet" href="//fonts.googleapis.com/css?family=Lato" />

Library: Requests

Author: Sarah Strochak

Version: R 3.6.1

Last updated `r format(Sys.time(), '%B %d, %Y')`

Environment: Local Windows session (desktop)

```{r rmarkdown-setup, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r setup}
options(scipen = 999)

library(tidyverse)
```

## Read in church-owned parcels

```{r read-churches}

churches <- read_csv("L:/Libraries/Requests/Raw/2019/dmv_who_owns_church.csv") %>% 
  mutate(fipscodestatecounty = as.character(fipscodestatecounty))

# Merge on county names
churches <- urbnmapr::get_urbn_map(map = "counties", sf = TRUE) %>% 
  sf::st_drop_geometry() %>% 
  select(county_name, county_fips) %>% 
  right_join(churches, by = c("county_fips" = "fipscodestatecounty"))

```

## Test building area and lot size variables.

```{r}

churches %>% 
  mutate(residential = ifelse(str_sub(zoning, 1, 1) == "R",
                              1,
                              0)) %>% 
  group_by(county_name, residential) %>% 
  summarize(parcels = n(),
            lotsize_missing = sum(is.na(lotsizeorarea)),
            lotsize_0 = sum(lotsizeorarea == 0),
            building_missing = sum(is.na(buildingarea)),
            building_0 = sum(buildingarea == 0))


```

## Create variables

1. Create vacant flag (detect county land use decription for "VACANT", "PARKING", "VAC")
2. Create residential flag (zones that are primarily residential- start with "R")
3. Recalculate lot area in square feet to avoud Black Knight conversion error.
4. Calculate units per lot. Assumptions are that you need 950 square feet per unit, and a lot covereage of 75 percent.

```{r}


churches1 <- churches %>% 
  mutate(buildingarea = ifelse(is.na(buildingarea),
                               0,
                               buildingarea),
         vacant = ifelse(str_detect(countylandusedescription, "VACANT") == TRUE,
                         1,
                         0),
         vacant = ifelse(str_detect(countylandusedescription, "PARKING") == TRUE,
                         1,
                         vacant),
         vacant = ifelse(str_detect(countylandusedescription, "VAC") == TRUE,
                         1,
                         vacant),
         vacant = ifelse(countylandusedescription %in%
                          c("FALSE VACANT",
                            "FALSE VACANT ABUTTING LOT (RES)"),
                         0,
                         vacant)) %>% 
  mutate(lotsize_sqfeet = ifelse(lotsizeareaunit == "SF",
                                 lotsizeorarea,
                                 lotsizeorarea * 43560),
         residential = ifelse(str_sub(zoning, 1, 1) == "R",
                              "Residential zone",
                              "Non-residential zone"),
         residential = ifelse(is.na(zoning),
                              "Missing zoning information",
                              residential),
         units_base = floor((lotsize_sqfeet * .75) / 950),
         units_2fl = units_base * 2,
         units_5fl = units_base * 5)


## Create extract for city of DC

churches_export <- churches1 %>% 
  filter(county_fips == "11001") %>% 
  mutate(vacant_status = ifelse(vacant == 0,
                                "Not vacant",
                                "Vacant")) %>% 
  mutate_at(vars(starts_with("units")),
         ~ ifelse(vacant == 0 | buildingarea > 0,
                  NA,
                  .)) %>% 
  select(county_name, county_fips,
         assessorsparcelnumberapnpin,
         starts_with("property"),
         -propertyaddresssourceflag,
         assesseeownername,
         countylandusedescription,
         zoning,
         lotsize_sqfeet,
         buildingarea,
         ownername_full,
         Ownercat,
         residential,
         vacant_status,
         starts_with("units"))

write_csv(churches_export,
          here::here("Prog", "2019", "church-parcels-units.csv"),
          na = "")

```


## Vacant parcels

Calculate units for 2 and 5 buildings, using only vacant and empty parcels.

```{r}

calc_units_vacant <- function(sq_feet = 950) {
  
  
  total_area <- churches1 %>% 
    group_by(county_name) %>% 
    summarize(total_parcels = n(),
              total_area_sqft = sum(lotsize_sqfeet, na.rm = TRUE),
              total_area_acres = total_area_sqft / 43560) %>% 
    mutate(residential = "Total") %>% 
    bind_rows(churches1 %>% 
                group_by(county_name, residential) %>% 
                summarize(total_parcels = n(),
                          total_area_sqft = sum(lotsize_sqfeet, na.rm = TRUE),
                          total_area_acres = total_area_sqft / 43560))
  
  totals <- churches1 %>% 
    filter(vacant == 1) %>% 
    mutate(units = units_base) %>% 
    group_by(county_name) %>% 
    summarize(vacant_parcels = n(),
              empty_parcels = sum(buildingarea == 0, na.rm = TRUE),
              total_empty_lot_sqft = sum(lotsize_sqfeet[buildingarea == 0], na.rm = TRUE),
              total_empty_lot_acres = total_empty_lot_sqft / 43560,
              average_empty_sqft = mean(lotsize_sqfeet[buildingarea == 0], na.rm = TRUE),
              average_empty_acres = average_empty_sqft / 43560,
              avg_units_2fl = mean(units[buildingarea == 0], na.rm = TRUE) * 2,
              total_units_2fl = sum(units[buildingarea == 0], na.rm = TRUE) * 2,
              avg_units_5fl = mean(units[buildingarea == 0], na.rm = TRUE) * 5,
              total_units_5fl = sum(units[buildingarea == 0], na.rm = TRUE) * 5) %>% 
    mutate(residential = "Total")
  
  
  all <- churches1 %>% 
    filter(vacant == 1) %>% 
    mutate(units = units_base) %>% 
    group_by(county_name, residential) %>% 
    summarize(vacant_parcels = n(),
              empty_parcels = sum(buildingarea == 0, na.rm = TRUE),
              total_empty_lot_sqft = sum(lotsize_sqfeet[buildingarea == 0], na.rm = TRUE),
              total_empty_lot_acres = total_empty_lot_sqft / 43560,
              average_empty_sqft = mean(lotsize_sqfeet[buildingarea == 0], na.rm = TRUE),
              average_empty_acres = average_empty_sqft / 43560,
              avg_units_2fl = mean(units[buildingarea == 0] * 2, na.rm = TRUE),
              total_units_2fl = sum(units[buildingarea == 0] * 2, na.rm = TRUE),
              avg_units_5fl = mean(units[buildingarea == 0], na.rm = TRUE) * 5,
              total_units_5fl = sum(units[buildingarea == 0], na.rm = TRUE) * 5) %>% 
    bind_rows(totals) %>% 
    left_join(total_area, by = c("residential", "county_name")) %>% 
    arrange(desc(residential), county_name) %>% 
    select(county_name, residential, total_parcels, total_area_sqft, total_area_acres,
           everything())
  
  
  write_csv(all, here::here("Prog", "2019", 
                            "church-lots_options_empty_vacant.csv"))
  
  
}

calc_units_vacant()


```
