---
title: "RFK Block Geography"
author: "Katie Visalli"
date: "2026-05-06"
output: html_document
---
Select census tracts that are within a mile of the RFK Stadium 
Lot 805 in Square 1128

Data From:
Square Shapefile: https://arcg.is/1TeO1b5
Census Tract Shapefile: https://arcg.is/1LjHaH2
RFK Stadium Parcel: https://www.ncpc.gov/files/projects/2026/8721_New_Stadium_at_RFK_Campus_Submission_Materials_Feb2026.pdf Page 5


```{r}
library(tidyverse)
library(sf)
library(ggplot2)
library(urbnthemes)
library(urbnmapr)
set_urbn_defaults(style = "map")
library(ggspatial)

install.packages("ggspatial")
```

```{r load data}
squares <- st_read("//sas1/dcdata/Libraries/Requests/Prog/2026/RFK/Data/Square_Boundaries/Square_Boundaries.shp")

tracts <- st_read("//sas1/dcdata/Libraries/Requests/Prog/2026/RFK/Data/Census_Tracts_in_2020/Census_Tracts_in_2020.shp")

wards <- st_read("//sas1/dcdata/Libraries/Requests/Prog/2026/RFK/Data/Ward_-_2022/Ward_-_2022.shp")

dc <- wards %>%
  st_combine()

anc <- st_read("//sas1/dcdata/Libraries/OCTO/Maps/Advisory_Neighborhood_Commissions_from_2023.shp")

roads <- st_read("//sas1/dcdata/Libraries/Requests/Prog/2026/RFK/Data/Roads/Roads.shp")

metro <- st_read("//sas1/dcdata/Libraries/Requests/Prog/2026/RFK/Data/Metro_Stations_Regional/Metro_Stations_Regional.shp") %>%
  mutate(line1 = case_when(LINE == "blue, yellow"~ "yellow",
                           LINE == "blue, orange, silver"~ "blue",
                           LINE == "blue, silver"~ "blue",
                           LINE == "green, yellow, orange, blue, silver"~ "green",
                           LINE == "green, yellow"~ "green",
                           LINE == "red, green, yellow" ~ "green",
                           LINE == "red, blue, orange, silver"~ "red", 
                           LINE == "red, green" ~ "green", 
                           LINE == "orange, silver" ~ "orange"
                           ),
         line_name = if_else(is.na(line1), LINE, line1), 
         line_color = if_else(line_name == "silver", "grey", line_name))
dc_metro <- metro %>%
  st_filter(., dc, .predicate = st_intersects)

bus <- st_read("//sas1/dcdata/Libraries/Requests/Prog/2026/RFK/Data/Metro_Bus_Stops_2023_-_2025/Metro_Bus_Stops_2023_-_2025.shp")

national_parks <- st_read("//sas1/dcdata/Libraries/Requests/Prog/2026/RFK/Data/National_Parks/National_Parks.shp") %>%
  mutate(parks = "National Parks")

build_permits <- st_read("//sas1/dcdata/Libraries/Requests/Prog/2026/RFK/Data/Building_Permits_in_2025/Building_Permits_in_2025.shp")
rfk_permit <- build_permits %>%
  filter(OBJECTID == 1097829048) %>%
  #filter(FULL_ADDRE == "2400 EAST CAPITOL ST SE, WASHINGTON, DC 20003") %>%
  mutate(label = "RFK Stadium Development")

```

```{r test maps}
map_squares <- ggplot() + 
  geom_sf(data = squares, aes(geometry = geometry), colour = "black", fill = NA)
map_squares

map_tracts <- ggplot() + 
  geom_sf(data = tracts, aes(geometry = geometry), colour = "red", fill = NA)
map_tracts

map_anc <- ggplot() + 
  geom_sf(data = anc, aes(geometry = geometry), colour = "black", fill = NA)
map_anc

map_roads <- ggplot() + 
  geom_sf(data = roads, aes(geometry = geometry), colour = "grey", fill = NA)
map_roads

map_metro <- ggplot() + 
  geom_sf(data = metro, aes(geometry = geometry), colour = "black", fill = NA)
map_metro

map_bus <- ggplot() + 
  geom_sf(data = bus, aes(geometry = geometry), colour = "grey", fill = NA)
map_bus

map_parks <- ggplot() + 
  geom_sf(data = national_parks, aes(geometry = geometry), colour = NA, fill = "green")
map_parks

map_wards <- ggplot() + 
  geom_sf(data = wards, aes(geometry = geometry), colour = "black", fill = NA)
map_wards

```
Find Lot 805 in Square 1128 in DC Parcel Data

There are two squares named 1128 in the data. One was updated in 2025 and has the same outline shape alone as the outline on page 5 of the concept review. So using it as my RFK square and ignoring the other, which is right next to, but much smaller than the 2025 version of 1128.

```{r find stadium}
squares2 <- squares %>%
  mutate(rfk = if_else(SQUARE == "1128" & OBJECTID == "1744567", 1, 0))

map_squares <- ggplot() + 
  geom_sf(data = squares2, aes(geometry = geometry, fill = rfk), color = NA)
map_squares

rfk <- squares2 %>%
  filter(rfk == 1) %>%
  mutate(label = "RFK Stadium & Campus")
```

Find shortest distance from each census tract to the RFK square
```{r}
tract2 <- tracts %>%
  mutate(distance_to_rfk = as.numeric(st_distance(., rfk))) %>%
  mutate(in_a_mile = if_else(distance_to_rfk <= 1609.34, 1, 0)) %>%
# Add census tracts 7803, 7806, and 7809 (Minnesota Ave Metro), Kennilworth and Eastwood gardens
# and remove the arboretum, 11100
  mutate(rfk_group = if_else(in_a_mile == 1, 1, 0), 
         rfk_group = if_else(TRACT %in% c("007803", "007806", "007809", "009906"), 1, rfk_group), 
         rfk_group = if_else(TRACT == "011100", 0, rfk_group), 
         east_of_anacostia = if_else(TRACT %in% c("007709", "007708", "007601", "007604", "009901", "007703", "009604", "009603", "009602", "007803", "009601", "007806", "007809", "009906"), 1, 0), 
         west_of_anacostia = if_else(rfk_group == 1 & east_of_anacostia == 0, 1, 0), 
         rfk_sub_group = case_when(east_of_anacostia == 1 ~ "RFK-East of Anacostia", 
                                   west_of_anacostia == 1 ~ "RFK-West of Anacostia"), 
         rfk_sub_group2 = case_when(east_of_anacostia == 1 ~ "East of Anacostia", 
                                   west_of_anacostia == 1 ~ "West of Anacostia"))

rfk_tracts <- tract2 %>%
  filter(rfk_group == 1)

rfk_outline <- rfk_tracts %>%
  st_union()

rfk_save <- tract2 %>%
  select(GEOID, rfk_group, east_of_anacostia, west_of_anacostia, rfk_sub_group, rfk_sub_group2) %>%
  st_drop_geometry()

write.csv(rfk_save, "//Sas1/dcdata/Libraries/Requests/Prog/2026/RFK/result/rfk_tracts.csv")
```

map result
```{r}
map_result <- ggplot() + 
  geom_sf(data = rfk_tracts, aes(geometry = geometry, fill = rfk_sub_group), colour = "white") + 
  geom_sf(data = rfk, aes(geometry = geometry), colour = "red", fill = "red")
map_result

# ggsave(filename = "//Sas1/dcdata/Libraries/RFK/Requests/Prog/2026/RFK/result/rfk_tract_map.png",
#        plot = map_result,
#        device = "png",
#        width = 6.5,
#        dpi = 1000)

```

```{r zoom-in area}

# Filter to Ward 7
ward7 <- wards %>%
  filter(WARD_ID == "7")

ward6 <- wards %>%
  filter(WARD_ID == "6")

#Find ward 7 coordinates, for zooming in on area
ward7coords = as.data.frame(unique(st_coordinates(ward7)))

#find rfk coordinates, for zooming in on area
rfk_coords = as.data.frame(unique(st_coordinates(rfk_tracts)))

#find x where y is max in ward 6 for x minnimum
ward6coords = as.data.frame(unique(st_coordinates(ward6))) %>%
  filter(Y == max(Y))


#minimmum of Y
ymin = min(ward7coords$Y)
ymax = max(ward7coords$Y)
xmin = ward6coords$X
xmax = max(ward7coords$X)

#DC Coordinates
dccoords = as.data.frame(unique(st_coordinates(wards)))

dcymin = min(dccoords$Y)
dcymax = max(dccoords$Y)
dcxmin = min(dccoords$X)
dcxmax = max(dccoords$X)

#Zoomed-in Metro Stops

```


```{r export map}
full_map <- ggplot() + 
  #RFK focus area
  geom_sf(data = rfk_tracts, aes(geometry = geometry, fill = rfk_sub_group), colour = NA) + 
  #National Parks
  geom_sf(data = national_parks, aes(geometry = geometry, fill = parks), colour = NA, alpha = 0.5) +
  #RFK Development
  geom_sf(data = rfk, aes(geometry = geometry, fill = label), colour = NA)  +
  #Wards
  geom_sf(data = wards, aes(geometry = geometry), colour = "#5c5859", fill = NA)  +
  #Metro Stops
  geom_sf(data = dc_metro, aes(geometry = geometry, colour = line_color), fill = NA, size = 2) +
  scale_color_identity(guide = "legend", name = "Metro Stops")+ 
  #Area legend
  scale_fill_manual(breaks = c("RFK Stadium & Campus", "RFK-West of Anacostia", "RFK-East of Anacostia", "National Parks"), values = c("#000000", "#fdbf11", "#ec008b", "#98cf90"), name = "Legend") +
  #Ward names
  geom_sf_text(data = wards, aes(geometry = geometry, label = paste0("Ward ", WARD_ID)), colour = "#5c5859", size = 3) + 
  #RFK area outline
  geom_sf(data = rfk_outline, color = "#000000", size = 3, fill = NA) +
  #Filter to DC
  coord_sf(
    xlim = c(dcxmin, dcxmax), 
    ylim = c(dcymin, dcymax),
    expand = FALSE) + 
  #distance label
  annotation_scale()
full_map

ggsave(filename = "//sas1/dcdata/Libraries/Requests/Prog/2026/RFK/Result/full_map.png",
       plot = full_map,
       device = "png",
       width = 6.5,
       dpi = 1000)

# 
# #Zoomed in Map with ANC
# full_map_zoom_anc <- ggplot() + 
#   #RFK focus area
#   geom_sf(data = rfk_tracts, aes(geometry = geometry, fill = rfk_sub_group), colour = NA) + 
#   #National Parks
#   geom_sf(data = national_parks, aes(geometry = geometry, fill = parks), colour = NA, alpha = 0.5) +
#   #RFK Development
#   geom_sf(data = rfk, aes(geometry = geometry, fill = label), colour = NA) + 
#   #Area legend
#   scale_fill_manual(breaks = c("RFK Stadium & Campus", "RFK-West of Anacostia", "RFK-East of Anacostia", "National Parks"), values = c("#000000", "#fdbf11", "#ec008b", "#98cf90"), name = "Legend") +
#   #ANC
#   geom_sf(data = anc, aes(geometry = geometry), colour = "#5c5859", fill = NA) +
#   geom_sf_text(data = anc, aes(geometry = geometry, label = paste0(NAME)), colour = "#5c5859", size = 3) +
#   #Wards
#   geom_sf(data = wards, aes(geometry = geometry), colour = "#5c5859", fill = NA) +
#   geom_sf(data = rfk, aes(geometry = geometry, fill = label), colour = NA) + 
#   #Metro stops
#   geom_sf(data = dc_metro, aes(geometry = geometry, colour = line_color), fill = NA, size = 2) +
#   scale_color_identity(guide = "legend", name = "Metro Stops") + 
#   #RFK area outline
#   geom_sf(data = rfk_outline, color = "#000000", size = 2, fill = NA)  +
#   #Zoom in
#   coord_sf(
#     xlim = c(xmin, xmax), 
#     ylim = c(ymin, ymax),
#     expand = FALSE) + 
#   #distance label
#   annotation_scale(pad_x = unit(5, "cm"))
# full_map_zoom_anc
# 
# 
# ggsave(filename = "//sas1/dcdata/Libraries/Requests/Prog/2026/RFK/Result/full_map_zoom_anc.png",
#        plot = full_map_zoom_anc,
#        device = "png",
#        width = 6.5,
#        dpi = 1000)


#Zoomed in Map with census tracts
full_map_zoom_tract <- ggplot() + 
  #RFK focus area
  geom_sf(data = rfk_tracts, aes(geometry = geometry, fill = rfk_sub_group), colour = NA) + 
  #National Parks
  geom_sf(data = national_parks, aes(geometry = geometry, fill = parks), colour = NA, alpha = 0.5) +
  #RFK Development
  geom_sf(data = rfk, aes(geometry = geometry, fill = label), colour = NA) + 
  #Area legend
  scale_fill_manual(breaks = c("RFK Stadium & Campus", "RFK-West of Anacostia", "RFK-East of Anacostia", "National Parks"), values = c("#000000", "#fdbf11", "#ec008b", "#98cf90"), name = "Legend") +
  #Census Tracts
  geom_sf(data = tract2, aes(geometry = geometry), fill = NA, colour = "#5c5859") +
  #Ward
  geom_sf(data = wards, aes(geometry = geometry), colour = "#5c5859", fill = NA) +
  #Metro
  geom_sf(data = dc_metro, aes(geometry = geometry, colour = line_color), fill = NA, size = 2) +
  scale_color_identity(guide = "legend", name = "Metro Stops") + 
  #RFK area outline
  geom_sf(data = rfk_outline, color = "#000000", size = 3, fill = NA)  +
  #Zoom in
  coord_sf(
    xlim = c(xmin, xmax), 
    ylim = c(ymin, ymax),
    expand = FALSE) + 
  #distance label
  annotation_scale(pad_x = unit(5, "cm"))
full_map_zoom_tract


ggsave(filename = "//sas1/dcdata/Libraries/Requests/Prog/2026/RFK/Result/full_map_zoom_tract.png",
       plot = full_map_zoom_tract,
       device = "png",
       width = 6.5,
       dpi = 1000)

#Zoomed in map with roads
full_map_zoom_roads <- ggplot() + 
  #RFK focus area
  geom_sf(data = rfk_tracts, aes(geometry = geometry, fill = rfk_sub_group), colour = NA) + 
  #National Parks
  geom_sf(data = national_parks, aes(geometry = geometry, fill = parks), colour = NA, alpha = 0.5) +
  #RFK Development
  geom_sf(data = rfk, aes(geometry = geometry, fill = label), colour = NA) + 
  #Area legend
  scale_fill_manual(breaks = c("RFK Stadium & Campus", "RFK-West of Anacostia", "RFK-East of Anacostia", "National Parks"), values = c("#000000", "#fdbf11", "#ec008b", "#98cf90"), name = "Legend") +
  #Roads
  geom_sf(data = roads, aes(geometry = geometry), colour = "#5c5859", fill = NA, size = 0.05) +
  # #Wards
  # geom_sf(data = wards, aes(geometry = geometry), colour = "#000000", fill = NA) + 
  #Metro
  geom_sf(data = dc_metro, aes(geometry = geometry, colour = line_color), fill = NA, size = 2) +
  scale_color_identity(guide = "legend", name = "Metro Stops") + 
  #RFK area outline
  geom_sf(data = rfk_outline, color = "#000000", size = 3, fill = NA) +
  #Zoom in
  coord_sf(
    xlim = c(xmin, xmax), 
    ylim = c(ymin, ymax),
    expand = FALSE) + 
  #distance label
  annotation_scale(pad_x = unit(5, "cm"))
full_map_zoom_roads


ggsave(filename = "//sas1/dcdata/Libraries/Requests/Prog/2026/RFK/Result/full_map_zoom_roads.png",
       plot = full_map_zoom_roads,
       device = "png",
       width = 6.5,
       dpi = 1000)


```


export results
```{r}
tract_list <- tract2 %>%
  select(TRACT, GEOID, rfk_group, rfk_sub_group) %>%
  st_drop_geometry()

write.csv(tract_list, "//sas1/dcdata/Libraries/Requests/Prog/2026/RFK/Result/rfk_distance_tracts.csv")
```

