---
title: "RFK Block Geography"
author: "Katie Visalli"
date: "2026-05-06"
output: html_document
---
Select census tracts that are within a mile of the RFK Stadium 
Lot 805 in Square 1128

Data From:
Square Shapefile: https://arcg.is/1TeO1b5
Census Tract Shapefile: https://arcg.is/1LjHaH2
RFK Stadium Parcel: https://www.ncpc.gov/files/projects/2026/8721_New_Stadium_at_RFK_Campus_Submission_Materials_Feb2026.pdf Page 5


```{r}
library(tidyverse)
library(sf)
library(ggplot2)
```

```{r}
squares <- st_read("//Sas1/dcdata/Libraries/RFK/Requests/Prog/2026/RFK/Data/Square_Boundaries.shp")

tracts <- st_read("//Sas1/dcdata/Libraries/RFK/Requests/Prog/2026/RFK/Data/Census_Tracts_in_2020.shp")
```
```{r}
map_squares <- ggplot() + 
  geom_sf(data = squares, aes(geometry = geometry), colour = "black", fill = NA)
map_squares

map_tracts <- ggplot() + 
  geom_sf(data = tracts, aes(geometry = geometry), colour = "red", fill = NA)
map_tracts

map_both <- ggplot() + 
  geom_sf(data = tracts, aes(geometry = geometry), colour = "red", fill = NA) + 
  geom_sf(data = squares, aes(geometry = geometry), colour = "black", fill = NA)
map_both

```
Find Lot 805 in Square 1128 in DC Parcel Data

There are two squares named 1128 in the data. One was updated in 2025 and has the same outline shape alone as the outline on page 5 of the concept review. So using it as my RFK square and ignoring the other, which is right next to, but much smaller than the 2025 version of 1128.

```{r}
squares2 <- squares %>%
  mutate(rfk = if_else(SQUARE == "1128" & OBJECTID == "1744567", 1, 0))

map_squares <- ggplot() + 
  geom_sf(data = squares2, aes(geometry = geometry, fill = rfk), color = NA)
map_squares

rfk <- squares2 %>%
  filter(rfk == 1)
```

Find shortest distance from each census tract to the RFK square
```{r}
tract2 <- tracts %>%
  mutate(distance_to_rfk = as.numeric(st_distance(., rfk))) %>%
  mutate(in_a_mile = if_else(distance_to_rfk <= 1609.34, 1, 0))
```

map result
```{r}
map_result <- ggplot() + 
  geom_sf(data = tract2, aes(geometry = geometry, fill = in_a_mile), colour = "white") + 
  geom_sf(data = rfk, aes(geometry = geometry), colour = "red", fill = "red")
map_result

ggsave(filename = "//Sas1/dcdata/Libraries/RFK/Requests/Prog/2026/RFK/result/rfk_tract_map.png",
       plot = map_result,
       device = "png",
       width = 6.5,
       dpi = 1000)

```

export results
```{r}
tract_list <- tract2 %>%
  select(TRACT, GEOID, distance_to_rfk, in_a_mile) %>%
  st_drop_geometry()

write.csv(tract_list, "//Sas1/dcdata/Libraries/RFK/Requests/Prog/2026/RFK/result/rfk_distance_tracts.csv")
```

