---
title: "Properties in RFK Areas"
author: "Katie Visalli"
date: "2026-06-24"
output: html_document
---

Property ownership
Source: Realprop.Parcel_base_who_owns
Universe: All active [in_last_ownerpt=1] residential [ui_proptype=10,11,12,13,19] property parcels

Parcels by property type [ui_proptype] and owner type [ownercat]
Single family houses [ui_proptype=10] by owner-occupancy [owner_occ_sale]
Single family houses [ui_proptype=10] with senior homestead exemption [hstd_code=5]
Condominium units [ui_proptype=11] by owner-occupancy [owner_occ_sale]
Condominium units [ui_proptype=11] with senior homestead exemption [hstd_code=5]

```{r set up}
library(tidyverse)
library(tidylog)
library(sf)
```


```{r Load data}
property_data <- read.csv("//sas1/dcdata/Libraries/Requests/Prog/2026/RFK/Data/owner_data.csv")

square_shape <- st_read("//sas1/dcdata/Libraries/Requests/Prog/2026/RFK/Data/Square_Boundaries/Square_Boundaries.shp")

rfk_tracts <- read.csv("//Sas1/dcdata/Libraries/Requests/Prog/2026/RFK/result/rfk_tracts.csv")

tract_shape <- st_read("//sas1/dcdata/Libraries/Requests/Prog/2026/RFK/Data/Census_Tracts_in_2020/Census_Tracts_in_2020.shp")

ward_shape <- st_read("//sas1/dcdata/Libraries/Requests/Prog/2026/RFK/Data/Ward_-_2022/Ward_-_2022.shp")

rfk_shape <- tract_shape %>%
  mutate(GEOID=as.numeric(GEOID)) %>%
  select(GEOID) %>%
  left_join(rfk_tracts, by = "GEOID")

```

Use shapefiles of census tracts and 
